---
title: "workflow_outline"
format: html
---

# Overall aims
Develop a mechanistic model of human-human infectious disease transmission that will be largely disease agnostic but as specificity builds will resemble Ebola virus disease.

# Model description

The code implements a stochastic individual-based **SI** (susceptible–infected) model on a time-varying random contact network.

## State and indexing

Let individuals be indexed $i \in \{1,\dots,N\}$ (`individual_index`) and discrete time steps $t \in \{1,\dots,T\}$ (`n_time_steps`). Let $X_i^{(t)} \in \{0,1\}$ be the infection status of individual $i$ at time $t$ (`infection_status_over_time[t, i]`), with $1$ = infected and $0$ = susceptible. Infection is absorbing — there is no recovery.

The initial condition (`init_infection_status`) seeds a single infected individual:

$$X_i^{(1)} = \mathbb{1}[i = 1], \qquad \text{i.e. } X^{(1)} = (1,0,\dots,0).$$

## Contact process (`simulate_contact_pairs`)

At each time step $t$ a fresh contact set is drawn. With $n_c$ contact events per individual (`n_contact_events`) and $M = n_c N$ candidate events, draw independently, with replacement,

$$a_k \sim \mathrm{Unif}\{1,\dots,N\}, \quad b_k \sim \mathrm{Unif}\{1,\dots,N\}, \qquad k = 1,\dots,M,$$

and discard self-pairs. The contact set at time $t$ is

$$\mathcal{E}^{(t)} = \big\{ (a_k, b_k) : a_k \neq b_k,\ k = 1,\dots,M \big\}.$$

The set of distinct contacts of individual $i$ collapses both stored directions (`contacts_1`, `contacts_2`, then `unique`):

$$\mathcal{N}_i^{(t)} = \{\, b : (i,b) \in \mathcal{E}^{(t)} \,\} \;\cup\; \{\, a : (a,i) \in \mathcal{E}^{(t)} \,\}.$$

## Infection update (`infection_function`)

For a susceptible individual $i$ ($X_i^{(t)}=0$), each individual $j$ contributes an independent Bernoulli transmission trial with success probability $p\,X_j^{(t)}$ only if $j$ is a contact (the length-$N$ `possible_infections` vector passed to `rbinom`):

$$B_{ij}^{(t)} \sim \mathrm{Bernoulli}\!\Big( p \, X_j^{(t)} \, \mathbb{1}[\,j \in \mathcal{N}_i^{(t)}\,] \Big), \qquad p = \texttt{p\_infection}.$$

The individual becomes infected if any trial succeeds (`max(outcome)`), and infected individuals remain infected (`if(state) return(TRUE)`):

$$X_i^{(t+1)} = \max\!\Big( X_i^{(t)},\ \max_{j} B_{ij}^{(t)} \Big).$$

Marginalising over the Bernoulli trials, and letting $k_i^{(t)} = \big|\{\, j \in \mathcal{N}_i^{(t)} : X_j^{(t)} = 1 \,\}\big|$ be $i$'s number of infectious contacts,

$$
X_i^{(t+1)} \mid X^{(t)} \ \sim\
\begin{cases}
\delta_{1} & X_i^{(t)} = 1,\\[4pt]
\mathrm{Bernoulli}\!\Big(1 - (1-p)^{\,k_i^{(t)}}\Big) & X_i^{(t)} = 0.
\end{cases}
$$

The time loop runs $t = 1,\dots,T-1$, generating $X^{(2)},\dots,X^{(T)}$, each with its own independent contact draw $\mathcal{E}^{(t)}$.

# Implementation

Daily universal probabilities
```{r}
seed <- set.seed(20260604)
params <- list(
  individual_index = 1:10,
  n_contact_events = 2,
  init_infection_status = c(TRUE, rep(FALSE, 9)),
  p_infection = 0.2,
  n_time_steps = 10
)

```

Create contact pairs prior to the timesteps
```{r}

simulate_contact_pairs <- function(
    individual_index  = params$individual_index,
    n_contact_events = params$n_contact_events) {
  
  n_individuals <- length(individual_index) 
  sample_size_contact_events <- n_contact_events*n_individuals

first_contact <- sample(x = individual_index, size = sample_size_contact_events, replace = TRUE)
second_contact <- sample(x = individual_index, size = sample_size_contact_events, replace = TRUE)

keep <- first_contact != second_contact 

# at the moment this allows for multiple contact events between the same pair in the same timestep
contact_pairs <- data.frame(first_contact = first_contact[keep], second_contact = second_contact[keep])

return(contact_pairs)
}

```

```{r}
# For testing
index <- 3

infection_function <- function(
    index,
    prob_infection,
    contact_pairs,
    infection_status
  ){
  state <- infection_status[index]
  
  if(state) {
    return(TRUE)
  } else {
    # contacts <- sample(
    #   x = infection_status[-index],
    #   size = n_contacts,
    #   replace = FALSE
    # )
  contacts_1 <- contact_pairs$second_contact[contact_pairs$first_contact == index]
  contacts_2 <- contact_pairs$first_contact[contact_pairs$second_contact == index]
  contacts <- c(contacts_1, contacts_2)
  contacts <- unique(contacts) # exclude multiple contact events between the same pair within the same timestep
    
  contacted <- seq_along(infection_status) %in% contacts
  infectious_contacts <- infection_status & contacted   
    
    possible_infections <- prob_infection*infectious_contacts
    
    outcome <- rbinom(
      n = length(possible_infections),
      size = 1,
      prob = possible_infections
    )
    
    return(as.logical(max(outcome)))
    
  }
}

```

```{r}
# Time-steps
# Currently takes in single contact_pairs object (should take in a unique one per time step)
infection_status_over_time <- matrix(nrow = params$n_time_steps, ncol = length(params$individual_index))
infection_status_over_time[1,] <- params$init_infection_status

contact_pairs_over_time <- list()

for (t in 1:((params$n_time_steps)-1)) {
  
  contact_pairs_over_time[[t]] <- simulate_contact_pairs()
  
  for (i in params$individual_index) {
    
    y <- infection_function(
      index = i, 
      prob_infection = params$p_infection, 
      contact_pairs = contact_pairs_over_time[[t]], 
      infection_status = infection_status_over_time[t,]
      )
    
    infection_status_over_time[t+1,i] <- y
  }
}

infection_status_over_time
# can print contact pairs for diagnostics
# print(contact_pairs_over_time)

```


