---
title: "workflow_outline"
format: html
---

# Overall aims
Develop a mechanistic model of human-human infectious disease transmission that will be largely disease agnostic but as specificity builds will resemble Ebola virus disease.

#

Daily universal probabilities
```{r}
seed <- set.seed(20260604)
# set timestep size in days
dt <- 1
params <- list(
  # simulation setup
  timestep_size = dt,
  individual_index = 1:10,
  init_infection_status = c(TRUE, rep(FALSE, 9)),
  n_time_steps = 10,
  # epi params
  n_contact_events_per_day = 2,
  p_infection_per_contact_event = 0.2,
  p_recovery_per_day = 0.1
)

### for next session: add compartments after infection (recovery and death)

```

Create contact pairs prior to the timesteps
```{r}

simulate_contact_pairs <- function(
    individual_index  = params$individual_index,
    n_contact_events_per_day = params$n_contact_events_per_day) {
  
  n_individuals <- length(individual_index) 
  sample_size_contact_events <- n_contact_events_per_day*n_individuals

first_contact <- sample(x = individual_index, size = sample_size_contact_events, replace = TRUE)
second_contact <- sample(x = individual_index, size = sample_size_contact_events, replace = TRUE)

keep <- first_contact != second_contact 

# at the moment this allows for multiple contact events between the same pair in the same timestep
contact_pairs <- data.frame(first_contact = first_contact[keep], second_contact = second_contact[keep])

return(contact_pairs)
}

```

```{r}
# For testing
index <- 3

infection_function <- function(
    index,
    prob_infection,
    contact_pairs,
    infection_status
  ){
  state <- infection_status[index]
  
  if(state) {
    return(TRUE)
  } else {
  contacts_1 <- contact_pairs$second_contact[contact_pairs$first_contact == index]
  contacts_2 <- contact_pairs$first_contact[contact_pairs$second_contact == index]
  contacts <- c(contacts_1, contacts_2)
    
  contacted <- seq_along(infection_status) %in% contacts # exclude multiple contact events between the same pair within the same timestep --- note we intend to look over this decision in the future
  infectious_contacts <- infection_status & contacted   
    
    possible_infections <- prob_infection*infectious_contacts
    
    outcome <- rbinom(
      n = length(possible_infections),
      size = 1,
      prob = possible_infections
    )
    
    return(as.logical(max(outcome)))
    
  }
}

```

Add a recovery process - add a per-timestep probability of becoming susceptible again (or later recovered, or partially recovered).

Options:
 1. start with a fixed per-timestep probability of moving to susceptible
 2. Have a dynamic process governing infection level
 

```{r}
recovery_function <- function(
    index,
    prob_recovery,
    infection_status
  ){
  
    this_one_is_infected <- infection_status[index]
    
    # if they are infected, make them susceptible with some probability
    if (this_one_is_infected) {
      # make them susceptible again with some probability
      # recovered <- as.logical(rbinom(1, 1, prob = prob_recovery))
      this_one_is_susceptible <- runif(1, min = 0, max = 1) < prob_recovery
    } else {
      this_one_is_susceptible <- TRUE
    }
    
    # return the new infection status
    this_one_is_infected <- !this_one_is_susceptible
    
    this_one_is_infected

}


```


```{r}
# Time-steps
# Currently takes in single contact_pairs object (should take in a unique one per time step)
infection_status_over_time <- matrix(nrow = params$n_time_steps, ncol = length(params$individual_index))
infection_status_over_time[1,] <- params$init_infection_status

contact_pairs_over_time <- list()

for (t in 1:((params$n_time_steps)-1)) {
  
  contact_pairs_over_time[[t]] <- simulate_contact_pairs()
  
  infected_after_infections <- rep(NA, length(params$individual_index))
  infected_after_recoveries <- rep(NA, length(params$individual_index))
  
  previously_infected <- infection_status_over_time[t,]

  
  # do infections
  for (i in params$individual_index) {
    
    y <- infection_function(
      index = i, 
      prob_infection = params$p_infection_per_contact_event, 
      contact_pairs = contact_pairs_over_time[[t]], 
      infection_status = previously_infected
      )
    
    infected_after_infections[i] <- y
  }
  
  # do recoveries
  for (i in params$individual_index) {

    # # binomial process ofr probability of recovery
    # p_not_recover_per_day <- 1 - params$p_recovery_per_day
    # n_days <- params$timestep_size
    # p_not_recover_per_timestep <- p_not_recover_per_day ^ n_days
    # p_recovery_per_timestep <- 1 - p_not_recover_per_timestep
    p_recovery_per_timestep <- 1 - (1 - params$p_recovery_per_day) ^ params$timestep_size
    
    # note in a compartmental type model with a rate, applied to a discrete timestep, the probability of changing state over that timestep is: 1 - exp(-rate * dt)

    
    y <- recovery_function(
      index = i, 
      prob_recovery = p_recovery_per_timestep, 
      # this doesn't include the new infections in this time step! That's nice?
      infection_status = previously_infected
    )
    
    infected_after_recoveries[i] <- y
  }

  newly_infected <- infected_after_infections & !previously_infected
  newly_recovered <- !infected_after_recoveries & previously_infected

  # what do we put in the new infection status?  
  currently_infected <- previously_infected & !newly_recovered |
    !previously_infected & newly_infected
  
  infection_status_over_time[t+1, ] <- currently_infected

}

infection_status_over_time
# can print contact pairs for diagnostics
# print(contact_pairs_over_time)

```
```{r}
image(infection_status_over_time)
```

