---
title: "workflow_outline"
format: html
---

# Overall aims
Develop a mechanistic model of human-human infectious disease transmission that will be largely disease agnostic but as specificity builds will resemble Ebola virus disease.

#

```{r}
set.seed(20260604)
individual_index <- 1:10


population_states <- c(TRUE, rep(FALSE, 9))

```


Daily universal probabilities
```{r}
# probability_of_infection_per_contact if contact is infected
p_infection <- 0.1

# number of contacts each day for each individual
n_contacts <- 2
```

# Model definition

Consider a closed population of $N = 10$ individuals, indexed by
$i \in \{1, \dots, N\}$, in discrete time (days) $t = 0, 1, 2, \dots$

**State.** Each individual has a binary infection state, collected in the
vector $\mathbf{X}^{(t)} = \big(X_1^{(t)}, \dots, X_N^{(t)}\big)$
(`population_states`):

$$
X_i^{(t)} \in \{0, 1\}, \qquad
X_i^{(t)} =
\begin{cases}
1 & \text{individual } i \text{ infected at time } t,\\
0 & \text{individual } i \text{ susceptible at time } t.
\end{cases}
$$

The initial condition seeds a single infection:

$$
X_1^{(0)} = 1, \qquad X_i^{(0)} = 0 \quad \text{for } i = 2, \dots, N.
$$

**Parameters.**

- $p \in [0, 1]$ — per-contact transmission probability from an infected
  contact ($p = 0.1$, `p_infection`);
- $k \in \mathbb{N}$ — number of contacts per individual per day
  ($k = 2$, `n_contacts`).

**Daily transition.** For a susceptible individual $i$, draw a contact set
uniformly at random, without replacement, from everyone else:

$$
C_i^{(t)} \subseteq \{1, \dots, N\} \setminus \{i\}, \qquad
\big|C_i^{(t)}\big| = k, \qquad
C_i^{(t)} \sim \text{Uniform}.
$$

Each drawn contact independently transmits with a probability proportional to
their infection state:

$$
Y_{ij}^{(t)} \,\big|\, \mathbf{X}^{(t)} \;\sim\;
\operatorname{Bernoulli}\!\big(p \, X_j^{(t)}\big),
\qquad j \in C_i^{(t)},
$$

so an uninfected contact ($X_j^{(t)} = 0$) can never transmit. The next state
is then (infection is absorbing — there is no recovery):

$$
X_i^{(t+1)} =
\begin{cases}
1, & X_i^{(t)} = 1,\\[6pt]
\displaystyle \max_{j \in C_i^{(t)}} Y_{ij}^{(t)}, & X_i^{(t)} = 0,
\end{cases}
$$

where the `max` over Bernoulli draws is equivalent to
$1 - \prod_{j \in C_i^{(t)}} \big(1 - Y_{ij}^{(t)}\big)$: individual $i$ becomes
infected if **at least one** sampled contact transmits.

**Implied one-step hazard.** Conditional on the contact set, let
$m_i^{(t)} = \sum_{j \in C_i^{(t)}} X_j^{(t)}$ be the number of infected
contacts. Then the probability that a susceptible individual is infected on the
next day is

$$
\Pr\!\big(X_i^{(t+1)} = 1 \,\big|\, \mathbf{X}^{(t)}, C_i^{(t)},\, X_i^{(t)} = 0\big)
= 1 - (1 - p)^{\,m_i^{(t)}}.
$$

Marginalising over the random contact draw, with
$I^{(t)} = \sum_{j \neq i} X_j^{(t)}$ the number of infected among the other
$N - 1$ individuals, $m_i^{(t)}$ follows a hypergeometric distribution and

$$
\Pr\!\big(X_i^{(t+1)} = 1 \,\big|\, \mathbf{X}^{(t)},\, X_i^{(t)} = 0\big)
= 1 - \sum_{m=0}^{k}
\frac{\dbinom{I^{(t)}}{m}\dbinom{N - 1 - I^{(t)}}{\,k - m\,}}{\dbinom{N - 1}{k}}
\,(1 - p)^{\,m}.
$$

The `infection_function` below implements a single application of this
transition for one individual $i$.

```{r}

infection_function <- function(
    index,
    p_infection,
    n_contacts,
    population_states
  ){
  state <- population_states[index]
  
  if(state) {
    return(TRUE)
  } else {
    contacts <- sample(
      x = population_states[-index],
      size = n_contacts,
      replace = FALSE
    )
    
    possible_infections <- p_infection*contacts
    
    outcome <- rbinom(
      n = length(possible_infections),
      size = 1,
      prob = possible_infections
    )
    
    return(as.logical(max(outcome)))
    
  }
}

```

```{r}

```


